Lightness constancy under fixed geometry

### 2017-10-12 Draft1/draft1.tex
edits

\title{Lightness constancy under fixed geometry}
\abstract{Abstract goes here.}

@@

In this work, we study a simpler analogue of color constancy called lightness constancy. Of the 3D human perceptual color space, we only consider the lightness component. Lightness is the attribute that determines the brightness of a color. A color with a low lightness is nearly black, while one with a high lightness is the pure, saturated color. For achromatic objects, it is the property that changes from black to gray to white. We study the effects of variations in the reflected light on the perceived lightness of an object.

@@

The variations in the reflected light can result from either geometrical factors (like shape, size, relative position of objects, etc.) or spectral factors (such as the illumination spectra, surface reflectance, diffusivity, etc.). In this work, we focus on the spectral factors. Specifically, we study the effects of variations in the surface reflectance spectra and the illumination spectra on lightness perception. We are interested in identifying the computations that can lead to stable lightness perception and constancy (Fig.\ref{fig:introExampleFigure}).

\begin{comment}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{approach.png}
\caption{{\bf Approach:} We have employed a supervised learning approach to study color constancy. We first generate a large dataset of multi-spectral images labeled with the lightness level of the target object. We generate images at multiple lightness levels, with images at the same lightness level having variable object reflectance spectra and illumination spectra. These multi-spectral images are used to simulate the response of the visual system. We use isetbio toolbox to model the optics and physiology of early vision and produce the response of the cones in the retina. We use supervised computational learning methods on this labeled dataset of cone-responses to estimate the lightness of the target object. The performance of the supervised method is calculated using a separate test dataset.} 
\label{fig:approach}
\end{figure}
\end{comment}

We employ a supervised learning approach to understand this phenomenon. Our approach involves generating a database of multi-spectral images of naturalistic scenes. These images are labeled by the lightness of a specific target object in the scene. The image label depends on the reflectance of the target object. It refers to the perceived lightness of the target object if it were placed under a standard light source ($D_{65}$ daylight spectrum). We generate hundreds of multi-spectral images at several fixed values of the target object standard lightness. At a particular lightness level, we allow variations in the surface reflectance spectra of the objects (both the target and the surround objects) and the illumination spectra of the light sources (Fig. \ref{fig:introExampleFigure}.

In this work, we study three cases resulting from the combination of such spectral variations (Fig.\ref{fig:studiedCases}. These cases are: 1. Variations in target object spectra (Fig.\ref{fig:targetVarying}) with fixed illumination and background surface reflectance spectra, 2. Variations in target object spectra and the light source spectra, with fixed background surface reflectance spectra (Fig.\ref{fig:targetIlluminantVarying}), and 3. Variations in all spectra (target object, background object and light source) (Fig.\ref{fig:allSpectraVarying}). The effects of other types of spectral variations can be explained in terms of one of these cases (see supplementary information).

@@

After generating databases of labeled multi-spectral images for each one of these cases, we simulate the response of the cones in the retina cone using an accurate model of the eye. These cone-responses datasets, that are labeled with the standard lightness of the target object, are then analyzed using supervised computational learning methods.

@@

We show that for spectral variations that allow variation either only in the target surface reflectance or the background surface reflectance or the spectral power distribution of the source of light, the standard lightness, $Y_{65}$ can be recovered by simple manipulations of the reflected light captured by the visual system. For example, if the illumination spectra of the light sources and the background are held fixed, while the target reflectance spectra change (case 1), $Y_{65}$ can be recovered directly from the light reflected off the target object. If the surface reflectance of the target object and the illumination spectra of the light sources are allowed to vary, but the background reflectance spectra is fixed (case 2), $Y_{65}$ can be recovered by from the contrast between the target and the background. When all the spectral features in the scene are allowed to change (case 3), $Y_{65}$ can not be recovered from a simple manipulations of the light captured by the visual system. Here, we have used a Bayesian supervised learning method called accuracy maximization analysis to estimate $Y_{65}$ of the target object. Using this method, we can recover the lightness of the target object with $\sim 15\%$ relative root mean square error.

Finally, we compared the generality of the methods for lightness estimation by comparing the optimal receptive fields of one case to the others. The optimal receptive fields for estimating the lightness of the target object show a center surround structure, supporting a comparison between the light coming from the target object and its surround. Additionally, the receptive fields give more weight to the L and M cone responses compared to S cone responses. We discuss the implications of such receptive fields on lightness constancy.

\section{Methods}
\subsection{Image database} \label{method:VirtualWorld}
In this section we briefly describe our image generation software, that we call Virtual world color constancy (VWCC). The details of the software are provided in the supplementary information section. To generate a virtual scene, we first select a 3D scene, a target object and a light source from the set of predefined scenes and objects of the VWCC software. The target object and the light source are inserted at predefined positions in the 3D scene. Fig.\ref{fig:3DScene} shows a 2D sRGB rendition of the 3D scene with the inserted target object (green sphere in the middle) used in our analysis. (The light source in not visible.) Next, we assign a surface reflectance spectrum and an illumination spectrum to each object and light source in the scene. The surface reflectance of the target object is assigned such that the perceived lightness of the target object under D65 daylight spectrum has a fixed value. Once these spectral properties have been assigned, we render a 320x240 pixel$^2$ 2d multi-spectral images of the scene at 31 frequencies (linearly spaced between 400nm - 700nm) using a graphics rendering system called Mitsuba \cite{jakob2015mitsuba}. We generate hundreds of such scenes and corresponding multi-spectral images for at 10 linearly spaced lightness levels between 0.2 and 0.6. For the images at a fixed lightness level, the reflectance and illumination spectra are assigned according to the spectral variations allowed under each case. Next, we crop a 51x51 pixel$^2$ part of the multi-spectral image containing the target object for further analysis (\ref{fig:croppedImage}). Fig. \ref{fig:studiedCases} shows examples of such cropped image at 3 lightness levels for the three cases considered in this work.

@@

\subsection{Model of Visual System} \label{method:Isetbio}
We simulate the visual system response using isetbio software toolbox. Isetbio simulates a model of the visual system accounting for cornea, pupil size, lens and physiological optics. We use the multi-spectral cropped image (fig.\ref{fig:croppedImage}) as the optical stimulus to simulate the response of the cones in the retina. The cropped image patch is assumed to cover a $1^{\circ}$ field of view at a distance of 1 m. The retinal cone mosaic is modeled as a square grid with 51x51 cones in the l:m:s ratio 0.6:0.3:0.1. The response of the cones are measured in terms of the number of isomerizations in the cone segment ({\it over what time ??}). Since, S cones response are much lower than of L and M cones, to model the amplification of the S cone response downstream, we rescale the cone response functions by the total area under the response curves. This leads to comparable number of isomerizations in the L, M and S cone on the presentation of an image. We interpolate the responses of cones in each cone class to get a $51 \times 51$ cone response image for each cone class.

@@

\subsection{Supervised Learning methods} \label{method:SupervisedLearning}
\subsubsection*{Linear Regression} To estimate the target object lightness using linear regression, we solve the equation $A_{(1,3)}*C_{(3,N)} = Y_{(1,N)}$. Here, $Y_{(1,N)}$ corresponds to the row vector of the lightness labels for the N images in the dataset. $C_{(3,N)}$ corresponds to the matrix of cone responses corresponding to the central pixel. Each column represents the response to one image and the rows corresponds to the contrast normalized L, M and S cone response of the center pixel. The coefficient matrix $A$ can be obtained as $A = Y*C^{-1}$.

The L, M and S response to the center pixel on the target object is calculate by taking the average of a 3x3 patch of cones near the center of the target.

@@

\begin{align}
N C^\intercal Y_D^2 C - 2(Y*\mathbbm{1})C^\intercal Y_D C + (Y*Y^\intercal) (C^\intercal C)
\end{align}

@@

To overcome this  challenge, we have developed an image rendering software pipeline that allows us to specify a virtual, but naturalistic, statistical model of visual scenes with parametric control over factors in natural scenes. The software uses a physically-based image rendering system that accurately accounts for the interactions of light with matter in 3D scenes. It then renders a multi-spectral 2D image of each modeled visual scene. Wherever possible, we have incorporated statistical models of the variation of specific scene parameters based on observed data in natural scenes. In addition, our software simulates the responses of retinal cones to the rendered multi-spectral images using a realistic model of the early human visual system. Below we provide the details of the software and the choices made to generate ensembles of scenes with corresponding images. Our software, which is written in MATLAB, is freely available through a public respository: \href{https://github.com/BrainardLab/VirtualWorldColorConstancy}{Virtual World Color Constancy}. It builds on several other open source projects, including \href{http://rendertoolbox.org}{RenderToolbox4} \cite{heasly2014rendertoolbox3}, \href{http://isetbio.org}{Isetbio} and \href{https://www.mitsuba-renderer.org}{Mitsuba} \cite{jakob2015mitsuba}. Below we refer to the software pipeline as VWCC (for Virtual World Color Constancy).

@@

After specifying the geometry and spectra, one of the inserted objects is chosen as the target object in the scene. A camera is inserted at a user specified position and then pointed at the target object. We ensure that the target object is not completely occluded by another object in the line of sight of the camera. Next, a predefined surface reflectance spectrum is assigned to the target object. This target surface reflectance sets the label for the virtual scene image. The scene is now rendered using the Mitsuba rendering package. The renderer produces 2D hyper-spectral images of the scene at specified wavelengths. Finally, these multi-spectral images are used to simulate retinal cone responses using the \href{https://github.com/isetbio}{ISETBIO} software.

@@

For the images used in this work, we first selected the ``{\it Library}'' base scene from the VWCC base scene set. We inserted the spherical ``{\it Big Ball}'' object in the base scene and called it the target object. We also inserted a light source with the same spherical ``Big Ball'' shape. Also, a camera is inserted the scene and pointed at the target object. The position, and size of the inserted object, light source and the camera was held fixed for all the images analyzed in this work, i.e., no geometrical manipulations were performed between the images. Next, we assigned illumination and surface reflectance spectra to the lights and objects in the scene. The spectral manipulations for the various cases studied in this work are described below. Multi-spectral 2D images were rendered at 31 wavelengths from $400$nm to $700$nm spaced $10$nm apart at an image size $320\times 240$ pixel$^2$. Next, a $51 \times 51$ pixel$^2$ patch of the multi-spectral image was cropped around the target object and cone responses were calculated to these cropped images as described below.

>>>>

\title{Lightness constancy under fixed geometry}
\abstract{Color constancy is the ability of human visual system to stably perceive the color of an object under variations in the light reflected from the object. This is a very significant invariance detection ability with important behavioral implications. The light captured by our visual system depends not just on the object of interest, but also the lighting and other objects in the surround. In order to corectly identify the object color, the variations resulting from the surround related factors must be discarded. To understand the computational principles that can lead to color constancy, we study constancy of the lightness component of the 3D color space. Lightness is the attribute of the color space that determines the brightness of the color. We study the effects of spectral variations, specifically the variations in the object reflectance spectra and the light source illumination spectra, on the perceived lightness of an object. We have employed a supervised learning approach to study lightness constancy. We have developed a computational software that generates naturalistic hyper-spectral images of 3D scenes. The software allows precise control of the geometrical and spectral properties of the objects and lights in the scene. It takes into account the statistical structure of the spectral variations in natural scenes. It also produces the response of the retinal cones to the hyperspectral images through an accurate model of human visual system. We label the iamges generated by the software through the standard lightness (lighness under standard daylight spectrum D65) of a specific target object in the scene. We study the corresponsding dataset of labeled cone responses using supervised computational learning methods. Our analysis shows that under simple spectral variations, like the variation only in the spectrum of the target object, or of the background object, or of the light source, the standard lightness can be recovered by simple trasnformations of the reflected light. When all spectra vary simultaneously, while it is not easy to recover the lightess through simple transformations, a decoding cheme that compares the light from the target and the surround and properly adds the response of the L,M and S cones can recover the lightness within about 15\% RMSE. }

@@

In this work, we study a simpler analogue of color constancy called lightness constancy. Lightness is the attribute that determines the brightness of a color. A color with a low lightness is nearly black, while one with a high lightness is the pure, saturated color. For achromatic objects, it is the property that changes from black to gray to white. Here, we only consider the lightness component of the  3D human perceptual color space and study the effects of variations in the reflected light on the perceived lightness.

@@

The variations in the reflected light can result from either geometrical factors (like shape, size, relative position of objects, etc.) or spectral factors (such as the illumination spectra, surface reflectance, diffusivity, etc.). In this work, we focus on the spectral factors. Specifically, we study the effects of variations in the surface reflectance spectra and the illumination spectra on lightness perception  (Fig.\ref{fig:introExampleFigure}). We are interested in identifying the computations that can lead to stable lightness perception.

\begin{comment}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{approach.png}
\caption{{\bf Approach:} We have employed a supervised learning approach to study color constancy. We first generate a large dataset of hyperspectral images labeled with the lightness level of the target object. We generate images at multiple lightness levels, with images at the same lightness level having variable object reflectance spectra and illumination spectra. These hyperspectral images are used to simulate the response of the visual system. We use isetbio toolbox to model the optics and physiology of early vision and produce the response of the cones in the retina. We use supervised computational learning methods on this labeled dataset of cone-responses to estimate the lightness of the target object. The performance of the supervised method is calculated using a separate test dataset.} 
\label{fig:approach}
\end{figure}
\end{comment}

We employ a supervised learning approach to understand this phenomenon. We generate a database of hyperspectral images of naturalistic scenes labeled by the lightness of a specific target object in the scene. The images are labeled with perceived lightness of a uniform surface having the target object reflectance placed under the standard daylight spectrum $D_{65}$. We generate hundreds of hyperspectral images at several fixed values of the target object standard lightness. At a particular lightness level, we allow variations in the surface reflectance spectra of the objects (both the target and the surround objects) and the illumination spectra of the light sources (Fig. \ref{fig:introExampleFigure}.

In this work, we study three cases resulting from the combination of such spectral variations (Fig.\ref{fig:studiedCases}). These cases are: 1. Variations in target object spectra (Fig.\ref{fig:targetVarying}) with fixed illumination and background surface reflectance spectra, 2. Variations in target object spectra and the light source spectra, with fixed background surface reflectance spectra (Fig.\ref{fig:targetIlluminantVarying}), and 3. Variations in all spectra (target object, background object and light source) (Fig.\ref{fig:allSpectraVarying}). The effects of other types of spectral variations can be explained in terms of one of these cases (see supplementary information).

@@

After generating databases of labeled hyperspectral images for each one of these cases, we simulate the response of the cones in the retina cone using an accurate model of the eye. These cone-responses datasets, that are labeled with the standard lightness of the target object, are then analyzed using supervised computational learning methods.

@@

We show that for cases that allow variation either only in the target surface reflectance or the background surface reflectance or the spectral power distribution of the source of light, the standard lightness ($Y_{65}$) can be recovered by simple manipulations of the reflected light captured by the visual system. For case 1, (variations only in target obejct reflectance spectra), $Y_{65}$ can be recovered directly from the light reflected off the target object. For case 2, where both the target object reflectance and light source spectra vary, $Y_{65}$ can be recovered by from the contrast between the target and the background. When all the spectral features in the scene are allowed to change (case 3), the standard lightness can not be recovered from a simple manipulations of the light captured by the visual system. Here, we have used two supervised learning methods with Bayesian estimators to estimate standard lightness of the target object. Using these methods, we can recover the lightness of the target object with $\sim 15\%$ relative root mean square error.

Our analysis shows that the optimal receptive fields for estimating the lightness of the target object show a center surround structure, supporting a comparison between the light coming from the target object and its surround. Additionally, lightness estimation invovles a weighted sum of the L, M and S cone responsese where the receptive fields give more weight to the L and M cone responses compared to S cone responses. Comparing the performace of the receptive fields of one case on the cone response of other cases shows that the receptive field of the complex cases perform well on images of simpler cases.  We discuss the implications of such receptive fields on lightness constancy.

\section{Methods}
\subsection{Image database} \label{method:VirtualWorld}
In this section we briefly describe our image generation software that we call Virtual world color constancy (VWCC). The details of the software are provided in the supplementary information section. To generate a virtual scene, we first select a 3D scene, a target object and a light source from the set of predefined scenes and objects of the VWCC software. The target object and the light source are inserted at predefined positions in the 3D scene. Fig.\ref{fig:3DScene} shows a 2D sRGB rendition of the 3D scene with the inserted target object (green sphere in the middle) used in our analysis. (The light source in not visible.) Next, we assign a surface reflectance spectrum and an illumination spectrum to each object and light source in the scene. For the target object the surface reflectance is assigned such that the perceived lightness of the target object under D65 daylight spectrum has a fixed value. Once these spectral properties have been assigned, we render a 320x240 pixel$^2$ 2d hyperspectral images of the scene at 31 frequencies (equal linear spacing between 400nm - 700nm) using a graphics rendering system called Mitsuba \cite{jakob2015mitsuba}. We generate hundreds of such scenes and the corresponding hyperspectral images at 10 linearly spaced lightness levels between 0.2 and 0.6. For the images at a fixed lightness level, the reflectance and illumination spectra are assigned according to the spectral variations allowed under each case. Next, we crop a 51x51 pixel$^2$ part of the hyperspectral image containing the target object for further analysis (\ref{fig:croppedImage}). Fig. \ref{fig:studiedCases} shows examples of such cropped image at 3 lightness levels for the three cases considered in this work.

@@

\subsection{Model of Visual System} \label{method:Isetbio}
We simulate the visual system response using \href{http://isetbio.org}{Isetbio} software toolbox. Isetbio simulates a model of the visual system accounting for cornea, pupil size, lens and physiological optics. We use the hyperspectral cropped image (fig.\ref{fig:croppedImage}) as the optical stimulus to simulate the response of the cones in the retina. The cropped image patch is assumed to cover a $1^{\circ}$ field of view at a distance of 1 m. The retinal cone mosaic is modeled as a square grid with 51x51 cones in the l:m:s ratio 0.6:0.3:0.1. The response of the cones are measured in terms of the number of isomerizations in the cone segment ({\it over what time ??}). Since, S cones response are much lower than of L and M cones, to model the amplification of the S cone response downstream, we rescale the cone response functions by the total area under the response curves. This leads to comparable number of isomerizations in the L, M and S cones. Further, to get the response of each cone class at each poistion on the retinal mosaic, we interpolate the response of the cones for each cone class. Thus we get a $51 \times 51$ cone response image for each cone class.

@@

\subsection{Supervised Learning methods} \label{method:SupervisedLearning}
\subsubsection*{Linear Regression} To estimate the target object lightness using linear regression, we solve the equation $A_{(1,3)}*C_{(3,N)} = Y_{(1,N)}$. $C_{(3,N)}$ corresponds to the matrix of LMS cone responses corresponding to the center pixel of the image. Each column represents the response to one image. The rows corresponds to the contrast normalized L, M and S cones response of the center pixel for the N images in the dataset. The (unknown) coefficient matrix $A$ can be obtained as $A = Y*C^{-1}$.

The L, M and S response to the center pixel on the target object is calculated by taking the average of a 3x3 patch of cones near the center of the target.

@@

\begin{align}
N C Y_D^2 C^\intercal - 2(Y*\mathbbm{1})C Y_D C^\intercal + (Y*Y^\intercal) (CC^\intercal)-C Y_{D}^2\mathbbm{1}\mathbbm{1}^\intercal C^\intercal +2 C Y_{D} \mathbbm{1}\mathbbm{1}^\intercal Y_{D} C^\intercal - C \mathbbm{1}\mathbbm{1}^\intercal Y_{D}^2 C^\intercal
\end{align}

@@

To overcome this  challenge, we have developed an image rendering software pipeline that allows us to specify a virtual, but naturalistic, statistical model of visual scenes with parametric control over factors in natural scenes. The software uses a physically-based image rendering system that accurately accounts for the interactions of light with matter in 3D scenes. It then renders a hyperspectral 2D image of each modeled visual scene. Wherever possible, we have incorporated statistical models of the variation of specific scene parameters based on observed data in natural scenes. In addition, our software simulates the responses of retinal cones to the rendered hyperspectral images using a realistic model of the early human visual system. Below we provide the details of the software and the choices made to generate ensembles of scenes with corresponding images. Our software, which is written in MATLAB, is freely available through a public respository: \href{https://github.com/BrainardLab/VirtualWorldColorConstancy}{Virtual World Color Constancy}. It builds on several other open source projects, including \href{http://rendertoolbox.org}{RenderToolbox4} \cite{heasly2014rendertoolbox3}, \href{http://isetbio.org}{Isetbio} and \href{https://www.mitsuba-renderer.org}{Mitsuba} \cite{jakob2015mitsuba}. Below we refer to the software pipeline as VWCC (for Virtual World Color Constancy).

@@

After specifying the geometry and spectra, one of the inserted objects is chosen as the target object in the scene. A camera is inserted at a user specified position and then pointed at the target object. We ensure that the target object is not completely occluded by another object in the line of sight of the camera. Next, a predefined surface reflectance spectrum is assigned to the target object. This target surface reflectance sets the label for the virtual scene image. The scene is now rendered using the Mitsuba rendering package. The renderer produces 2D hyper-spectral images of the scene at specified wavelengths. Finally, these hyperspectral images are used to simulate retinal cone responses using the \href{https://github.com/isetbio}{ISETBIO} software.

@@

For the images used in this work, we first selected the ``{\it Library}'' base scene from the VWCC base scene set. We inserted the spherical ``{\it Big Ball}'' object in the base scene and called it the target object. We also inserted a light source with the same spherical ``Big Ball'' shape. Also, a camera is inserted the scene and pointed at the target object. The position, and size of the inserted object, light source and the camera was held fixed for all the images analyzed in this work, i.e., no geometrical manipulations were performed between the images. Next, we assigned illumination and surface reflectance spectra to the lights and objects in the scene. The spectral manipulations for the various cases studied in this work are described below. hyperspectral 2D images were rendered at 31 wavelengths from $400$nm to $700$nm spaced $10$nm apart at an image size $320\times 240$ pixel$^2$. Next, a $51 \times 51$ pixel$^2$ patch of the hyperspectral image was cropped around the target object and cone responses were calculated to these cropped images as described below.
